NOPT048 Notes

### 2019-08-13 misc/study-notes/abstract-algebra.tex
Changed LaTeX to TeX
Added proofmac

\section 1.1. Basic properties of square matrices

@@

For two matrices $A$ and $B$, we may form their {\it sum} $A+B = (a_{ij} + b_{ij})$. (This notation means that at the $i$th row and $j$th column, the entry is the sum of $a_{ij}$ and $b_{ij}.)$. Given a real number $\alpha$, we obtain the scalar product $\alpha A  = (\alpha\cdot a_{ij})$ by multiplying every entry in $A$ by $\alpha$.

@@

\section 1.2. The general linear group

@@

\section 2.1. Properties and basic examples

@@

\section 2.3 Isomorphisms and homomorphisms

@@

>>>>

\section 1.1. Basic Properties of Square Matrices

@@

For two matrices $A$ and $B$, we may form their {\it sum} $A+B = (a_{ij} + b_{ij})$. (This notation means that at the $i$th row and $j$th column, the entry is the sum of $a_{ij}$ and $b_{ij}.)$ Given a real number $\alpha$, we obtain the scalar product $\alpha A  = (\alpha\cdot a_{ij})$ by multiplying every entry in $A$ by $\alpha$.

@@

\section 1.2. The General Linear Group

@@

\section 2.1. Properties and Basic Examples

@@

\section 2.3 Isomorphisms and Homomorphisms

@@

\section ?.?. Group Actions on Sets

Let $G$ be a group and $S$ a set of elements. We say that $G$ {\it acts on} $S$ if there exists a map from $G\times S\rightarrow S$ (whose pairs $(g,s)$ are written $g(s)$) such that, for all $s\in S$, $e(s) = s$ and $(gh)(s) = g\big(h(s)\big)$ for every $g,h\in G$. For any element $s\in S$, the {\it orbit} is set of elements in $S$ to which $s$ may be taken by $G$:
$$O_s = \{g(s)\in S : g\in G\}$$
If there exists an $s$ such that $O_s = S$, then we say that $G$ acts {\it transitively} on $S$. In fact, because the orbits of a group action partition the set $S$ into equivalence classes, transitivity of a group action implies there is only one orbit, and for any $s,s'\in S$, there exists some $g\in G$ such that $g(s) = s'$.

The {\it stabiliser} of $s$ is the set of all elements in $G$ that leave $s$ fixed:
$$G_s = \{g\in G : g(s) = s\}$$

Consider a model example of a transitive action. Let $G$ be a group with a subgroup $H\leq G$; let $S$ be the set $\{aH : a\in G\}$ of left cosets of $H$. The group $G$ acts on $S$ by taking a coset $aH$ to $gaH$. This action is transitive: For any $a, a'$, there exists some $g$ such that $g(a) = a'$, so we can take any $aH$ to $a'H$ via $g$ as well. The closure property of subgroups gives us $G_H = H$, and the stabiliser of an arbitrary coset $aH$ is the set $\{g\in G : gaH = aH\}$. We find that this is exactly the conjugate subgroup $aHa^{-1}$. More generally, if $G$ acts transitively on a set $S$ and $g(s) = s'$, then $G_{s'} = gG_sg^{-1}\subset G$.

\section ?.?. The Sylow Theorems

Suppose that a group $G$ acts on itself by conjugation:
$$ s\mapsto gsg^{-1}.$$
Then the orbit of an element $s$ is its conjugacy class and its stabiliser is the set $\{g\in G : gsg^{-1}\}$, also called the centraliser. Let $\overline H$ denote the set of all subgroups of $G$. Then $G$ also acts on $\overline H$ by conjugation: $g(H) = gHg^{-1}$. In this action, the stabiliser of a subgroup $H$ is the set
$$ G_H = \{ g \in G : gHg^{-1} = H\}.$$
Note that this is not the centraliser of $H$; elements may not be taken to themselves pointwise, so long as the sets $gHg^{-1}$ and $H$ are equal. This subset of $G$ is called the {\it normaliser} of $H$ and is itself a subgroup, denoted $N(H)$.
